---
title: "Project"
author: "Sumanth Paupuleti"
date: "2024-03-12"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#The dataset contains information about Connecticut Real Estate Sales Data, 
#including their typeSerial Number
```{r , echo=F , message=F,warning=F}
library(tidyverse)
library(dplyr)
library(tidyselect)
library(ggplot2)


#Data
sales_data <- read_csv(("C:\\Users\\srees\\Downloads\\sales_data.csv"))
```

QUESTION 1
To calculate the total number of sales
```{r , echo=F , message=F,warning=F}
total_sales<-sum(sales_data$`Sale Amount`)
total_sales
barplot(height = sales_data$`Sale Amount`,
        names.arg = rownames(sales_data),
        main = "total sales",
        xlab = "Category",
        ylab = "Sale Amount",
        col = "skyblue",
        ylim = c(0, max(sales_data$`Sale Amount`) * 1.1))
```

What are the top 5 towns with the highest total assessed value?
```{r , echo=F , message=F,warning=F}
Assesed_values <-sales_data %>% group_by(Town) %>% summarise(`Total Assessed Value`= sum(`Assessed Value`))
top_5<- Assesed_values %>% arrange(desc(`Total Assessed Value`)) %>% head(5)
ggplot(top_5, aes(x = reorder(Town, `Total Assessed Value`), y = `Total Assessed Value`)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Top 5 Towns with Highest Total Assessed Value",
       x = "Town",
       y = "Total Assessed Value") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(label = scales::comma(`Total Assessed Value`)), vjust = -0.3) 
```
QUESTION 3
identifying key trends in property values across Connecticut
```{r , echo=F , message=F,warning=F}
avg_sale_price_by_year <- sales_data %>%
  group_by(`List Year`) %>%
  summarize(mean_sale_price = mean(`Sale Amount`))
avg_sale_price_by_year

ggplot(avg_sale_price_by_year, aes(x = `List Year`, y = mean_sale_price)) +
  geom_line(color = "blue") +
  geom_point(color = "blue") +
  labs(title = "Average Sale Price by Year",
       x = "Year",
       y = "Average Sale Price") +
  theme_minimal()
```
3.1Calculate average sale price by property type
```{r , echo=F , message=F,warning=F}
avg_sale_price_by_property_type <- sales_data %>%
  group_by(`Property Type`) %>%
  summarize(mean_sale_price = mean(`Sale Amount`))
avg_sale_price_by_property_type

ggplot(avg_sale_price_by_property_type, aes(x = `Property Type`, y = mean_sale_price)) +
  geom_bar(stat = "identity", fill = "lightgreen") +
  labs(title = "Average Sale Price by Property Type",
       x = "Property Type",
       y = "Average Sale Price") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
QUESTION 4
Assess the relationship between assessed value and sale amount
```{r , echo=F , message=F,warning=F}
model_1 <- lm(`Sale Amount` ~ `Assessed Value`, data = sales_data)
model_1
ggplot(sales_data, aes(x = `Assessed Value`, y = `Sale Amount`)) +
  geom_point() +  # Plot the points
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE) +  # Add a linear regression line
  labs(title = "Relationship between Assessed Value and Sale Amount",
       x = "Assessed Value",
       y = "Sale Amount") +
  theme_minimal()
```
QUESTION 5
Explore the distribution of property types and their impact on values.
grouped_data_stats <- sales_data %>%
  group_by(`Property Type`, `Town`, `List Year`) %>%
  summarize(
    n = n(),
    mean_sale = mean(`Sale Amount`),
    median_sale = median(`Sale Amount`),
    min_sale = min(`Sale Amount`),
    max_sale = max(`Sale Amount`),
    sd_sale = sd(`Sale Amount`)
  )

```{r , echo=F , message=F,warning=F}
ggplot(sales_data, aes(x = `Sale Amount`, fill = `Property Type`)) +
  geom_histogram(binwidth = 50000, position = "identity", alpha = 0.7) +
  labs(title = "Distribution of Sale Amounts by Property Type",
       x = "Sale Amount",
       y = "Frequency") +
  theme_minimal() +
  facet_wrap(~`Property Type`, scales = "free_y")  # Separate facets for each property type
```


selected_property_type_1 <- "Residential"
property_types_1 <- sales_data[sales_data$`Property Type` == selected_property_type_1, ]



```{r , echo=F , message=F,warning=F}
selected_property_type_1 <- "Residential"
property_types_1 <- sales_data[sales_data$`Property Type` == selected_property_type_1,]
ggplot(property_types_1, aes(x = factor(`Property Type`), y = `Sale Amount`)) +
  geom_boxplot(fill = "lightgreen") +
  labs(title = paste("Boxplot of Sale Amounts for", selected_property_type_1),
       x = "Property Type",
       y = "Sale Amount") +
  theme_minimal()
```
QUESTION 6
Investigate factors influencing property values, including location, amenities, property type, and market conditions.
6.1 Average Sale Price by Year and Property Type
```{r , echo=F , message=F,warning=F}
grouped_data_stats <- sales_data %>%
  group_by(`Property Type`, `Town`, `List Year`) %>%
  summarize(
    n = n(),
    mean_sale = mean(`Sale Amount`),
    median_sale = median(`Sale Amount`),
    min_sale = min(`Sale Amount`),
    max_sale = max(`Sale Amount`),
    sd_sale = sd(`Sale Amount`)
  )
ggplot(grouped_data_stats, aes(x = `List Year`, y = mean_sale, color = `Property Type`)) +
  geom_line() +
  labs(title = "Average Sale Price by Year and Property Type", x = "Year", y = "Average Sale Price", color = "Property Type")
```

6.2 average sale price by town for different property types
```{r , echo=F , message=F,warning=F}
ggplot(grouped_data_stats, aes(x = Town, y = mean_sale, fill = `Property Type`)) +
  geom_violin()+
  labs(title = "Average Sale Price by Town and Property Type", x = "Town", y = "Average Sale Price", fill = "Property Type")
```


QUESTION 7
The  properties with assessed value is more than the  sale value

```{r , echo=F , message=F,warning=F}
properties_with_higher_assessed_value <- sales_data[sales_data$`Assessed Value` > sales_data$`Sale Amount`, ]
head(properties_with_higher_assessed_value)
barplot(height = properties_with_higher_assessed_value$`Assessed Value` - properties_with_higher_assessed_value$`Sale Amount`,
        names.arg = properties_with_higher_assessed_value$PropertyID,
        xlab = "Property ID",
        ylab = "Difference (Assessed Value - Sale Value)",
        main = "Difference between Assessed Value and Sale Value for Properties with Higher Assessed Value")
```




